Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-02 paper/structureScene.tex
Edit intro

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Because they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is that analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling} and needs some sort of pre-screening to reduce the need for human expert listening and annotation.

The most straightforward way is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}. By doing so, a period of time, say a day is represented as the number of events within the predefined typology that have been detected in this period of time. This approach allows the scientist to reduce drastically the amount of information that needs to be processed.  Though, this approach have two drawbacks. The first is the reliance to trained models whose behavior on unseen data (different sensors, ...) is prone to errors and can hardly be asserted. And, more importantly, it is based on \textit{a priori} knowledge, and thus cannot be considered for exploratory analysis where the quantities of interest are yet to be defined.

and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

To this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset. This requires the audio fragments to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches in the litterature\cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering decent size evaluation datasets \cite{lagrange:hal-01082501}.

@@

Addressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \ja{some relevant refs on dnns 4 audio}.

For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained features. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined. Thus leading to an unserpervised setting.

@@

%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. In Section \ref{sec:experiments}, several experiments are described for the acoustic scene similarity retrieval task. Results are subsequently reported in Section \ref{sec:results}.

>>>>

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.

This work falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.
An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}. Some sort of pre-screening is therefore needed to reduce the need for human expert listening and annotation.

The most straightforward approach is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}. A given time interval, for example a single day, is then represented by the number of events within the predefined typology that have been detected in that period of time. This approach allows the scientist to drastically reduce the amount of information that needs to be processed. However, this approach has two drawbacks. The first is the reliance to trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be always be trusted. More importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where the quantities of interest have yet to be defined.

\ml{What's going on here?}
and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

With this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset. This requires the audio fragments to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented with scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely human acoustic perception of acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be insufficient when considering evaluation datasets of typical sizes \cite{lagrange:hal-01082501}.

@@

Addressing this task first requires the ability to capture meaningful aspects within the signal at small time-scales.
In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \ja{some relevant refs on dnns 4 audio}.

For the task at at hand, scattering features have the advantage of not requiring training and thus potentially lead to wider range of application than trained features. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined, thus leading to an unserpervised setting.

@@

%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss feature post-processing in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task. Results are subsequently reported in Section \ref{sec:results}.
